---
output: 
    # github_document: default
    html_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  # cache = TRUE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```
# dcmisc

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/dcmisc)](https://cran.r-project.org/package=dcmisc)
<!-- badges: end -->

The goal of dcmisc is to tabulate and compare variables' values and their labels of labelled dataframes.

## Installation

Install the package via:

```{r, install}
# install.packages("K:/Tools/R/self-written packages/dcmisc/", type = "source", repos = NULL)
```

## Introduction

The package dcmisc mainly consists of 2 families of functions:

* `tab_*` tabulating count and label information of dataframes (denoted `df`), and
* `cmp_*` comparing this data for all dataframes in a list `l`.

The following attributes are tabulated / compared:

* `var` The name of the variables in the dataframe(s).
* `type` The type of the variables in the dataframe(s).
* `varlab` The variable label of the variable in the dataframe. 
* `val` The value of the variable in the dataframe. 
* `vallab` The value label of the variable in the dataframe. 

Comparisons further include `ex`, denoting whether the value is existent in the dataframe.

## Example


```{r data}
library(dcmisc)
library(tidyverse)


path <- system.file("examples", "iris.sav", package = "haven")
df <- haven::read_sav(path)
df <- tibble(id = 1:10,
       sex = haven::labelled(c(2, 1, 2, 1, 1, 2, 2, 1, 2, 1),
                             label = "sex",
                             labels = c(MALES = 1, FEMALES = 2)),
       age = c(24, 23, 23, 41, 23, 39, 30, 18, 31, 48),
       marital = haven::labelled(c(1, 7, 2, 6, 4, 5, 3, 8, 4, 2),
                                 label = "marital status",
                                 labels = c("single" = 1,
                                            "steady relationship" = 2,
                                            "living with partner" = 3,
                                            "married first time" = 4,
                                            "remarried" = 5,
                                            "separated" = 6,
                                            "divorced" = 7,
                                            "widowed" = 8))
       )
df
```


Create modified copy:

```{r, mod1}
df2 <- 
  df %>% 
  # Remove rows 9 & 10:
  slice(-c(9:10)) %>% 
  # Create Variable new_var equal to 1:
  mutate(new_var = 1)
# Change variable label:
attr(df2$new_var, "label")  <-  "new something"
```

Tabulate the variable labels for the dataframes in the list:

```{r, tab_varl}
list(df, df2) %>% map(tab_varlabs)
```


Change value:

```{r, mod2}
df2[1:2,"sex"] <- 3
```

Tabulate the count comparison:

```{r, tab_cts}
list(df, df2) %>% cmp_cts() %>% DT::datatable(class = "compact",                   
                                  options = list(scrollX = TRUE))

```

More label and value modifications:


```{r, more_mod}
attr(df2$sex, "labels")  <-  c(male = 1, female = 2, `Third Gender` = 3)
# Remove variable label:
attr(df2$sex, "label")  <-  NULL
# Change variable "marital" to 9 in rows 3 & 4:
df2[3:4,"marital"] <- 9
# Add another value label for the code 9:
attr(df2$marital, "labels") <- c(attr(df2$marital, "labels"), c("other" = 9))

df2
```

Tabulate the variable labels for the dataframes in the list:

```{r, tab_vall}
list(df, df2) %>% map(tab_vallabs)
```



Tabulate the comparison of count and label data and only show rows with differences:

```{r, anydiff}
list(df, df2) %>% 
  cmp_all() %>% 
  filter(any_diff) %>% 
  DT::datatable(class = "compact",                   
                options = list(scrollX = TRUE))
```

Show all cases (including a column with the ids), where the values in the first dataframe are equal to 1 and 2:

```{r, val1_1or2}
list(df, df2) %>% 
  cmp_all(include_ids = TRUE, 
          spec_diffs = c("val"), 
          col_groups = "spec") %>% 
  filter(val1 %in% 1:2) %>% 
  DT::datatable(class = "compact",                   
                options = list(scrollX = TRUE))
```


# Alte Sachen

## weitere infos

Hilfe anzeigen lassen über: ?tab_varlabs oder ?tab_vallabs oder durch
Mit-cursor-auf-eine-der-Funktionen-Gehen und F1 drücken (das geht aber nicht
im Kommentar) und nur nachdem die library schon geladen wurde

```{r}
  
# message, wenn keine value label im data.frame sind
tab_vallabs(data.frame(NA))

```

informative Fehler-Nachricht wenn df kein data.frame ist:

```{r}
# tab_vallabs(1)
```

erzeugt den Fehler:

```{r}
# Error: df is not a data frame
```


wenn Du anfängst eine der Funktionen zu schreiben, kommt dann auch eine kleine
informative Hilfe-Box für die geforderten Eingangs-Argumente, sobald die
autocomplete-Vorschäge erscheinen (so wie das Excel auch macht...)

