---
output: 
    github_document: default
    # html_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  # cache = TRUE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```
# tablab

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tablab)](https://cran.r-project.org/package=tablab)
<!-- badges: end -->

The goal of tablab is to tabulate and compare variables' values and their labels of labelled dataframes.

## Installation

Install the package via:

```{r, install, eval=FALSE}
devtools::install_github("urswilke/tablab")
```

## Introduction

The package tablab mainly consists of 2 families of functions:

* `tab_*` tabulating count and label information of dataframes (denoted `df`), and
* `cmp_*` comparing this data for all dataframes in a list `l`.

The following attributes are tabulated / compared:

* `var` The name of the variables in the dataframe(s).
* `type` The type of the variables in the dataframe(s).
* `varlab` The variable label of the variable in the dataframe. 
* `val` The value of the variable in the dataframe. 
* `vallab` The value label of the variable in the dataframe. 


## Examples

### Simple example
Assume you have a dataframe of labelled variables:

```{r}
library(tablab)
library(dplyr)
library(purrr)
set.seed(1)

x <- haven::labelled(
  sample(1:10), 
  label = "variable label", 
  labels = c("1lab" =1 , "2lab" = 2, "3lab" = 3)
)
y <- haven::labelled(
  LETTERS[sample(2:6, size = 10, replace = TRUE)], 
  label = "char_var_lab", 
  labels = c("Alab" ="A" , "Blab" = "B", "Clab" = "C")
)

df <- data.frame(id = 1:10, x, y)
df
```

tablab consists of functions to tabulate summaries of these labelled dataframes. See for instance the function:
```{r}
tab_all(df)
```

The result shows for every variable `var` in the data, its numeric/character values `nv`/`cv`, their counts `n`, as well as the variable and value labels `varlab` and `vallab`.

### More examples

```{r data}


path <- system.file("examples", "iris.sav", package = "haven")
df <- haven::read_sav(path)
df <- tibble(id = 1:10,
       sex = haven::labelled(c(2, 1, 2, 1, 1, 2, 2, 1, 2, 1),
                             label = "sex",
                             labels = c(MALES = 1, FEMALES = 2)),
       age = c(24, 23, 23, 41, 23, 39, 30, 18, 31, 48),
       marital = haven::labelled(c(1, 7, 2, 6, 4, 5, 3, 8, 4, 2),
                                 label = "marital status",
                                 labels = c("single" = 1,
                                            "steady relationship" = 2,
                                            "living with partner" = 3,
                                            "married first time" = 4,
                                            "remarried" = 5,
                                            "separated" = 6,
                                            "divorced" = 7,
                                            "widowed" = 8))
       )
df
```


Create modified copy:

```{r, mod1}
df2 <- 
  df %>% 
  # Remove rows 9 & 10:
  slice(-c(9:10)) %>% 
  # Create Variable new_var equal to 1:
  mutate(new_var = 1)
# Change variable label:
attr(df2$new_var, "label")  <-  "new something"
```

Tabulate the variable labels for the dataframes in the list:

```{r, tab_varl}
list(df, df2) %>% map(tab_varlabs)
```


Change value:

```{r, mod2}
df2[1:2,"sex"] <- 3
```

Tabulate the count comparison:

```{r, tab_cts}
list(df, df2) %>% cmp_cts()

```

More label and value modifications:


```{r, more_mod}
attr(df2$sex, "labels")  <-  c(male = 1, female = 2, `Third Gender` = 3)
# Remove variable label:
attr(df2$sex, "label")  <-  NULL
# Change variable "marital" to 9 in rows 3 & 4:
df2[3:4,"marital"] <- 9
# Add another value label for the code 9:
attr(df2$marital, "labels") <- c(attr(df2$marital, "labels"), c("other" = 9))

df2
```

Tabulate the variable labels for the dataframes in the list:

```{r, tab_vall}
list(df, df2) %>% map(tab_vallabs)
```



Tabulate the comparison of count and label data and only show rows with differences:

```{r, anydiff}
list(df, df2) %>% 
  cmp_all() %>% 
  filter(any_diff)
```

